---
title: "MA-217 Final Project"
author: "Thomas Gillis"
date: "2024-05-06"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r echo = FALSE, warning=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r echo = TRUE, warning=FALSE, message=FALSE, results='hide'}
library(tidyverse)

library(igraph)

library(ggraph)

library(latex2exp)

library(e1071)

```

## Loading Data Set

Data from Evelina Gabasova on
[Github](https://github.com/evelinag/StarWars-social-network/blob/129fbbdb8919e0dd14bc2db4a44f496a14e5ea07/networks/starwars-full-interactions-allCharacters-merged.json),
converted to `.graphml` and uploaded to Github. Other relevant files can
be found in the
[repo](https://github.com/tomdotgillis/MA217_Final_Project/tree/15260084fbb2d23533d993ad4f28cebead1b57be).

```{r echo = FALSE, warning=FALSE, results='hide'}
g <- read_graph('https://raw.githubusercontent.com/tomdotgillis/MA217_Final_Project/main/StarWarsGraph_evelinag.graphml', format = 'graphml')
```

## Loading Data from the Graph

```{r echo = FALSE, warning=FALSE, results='hide'}
# Vertices
V(g) # Vertices
E(g) # Edges
node_colors <- V(g)$colour # Colors assigned in the file
components(g)$no# Number of Components
components(g)$csize# Size of Components
THX <- 1138
```

**1. What is the order of your network? What is the size of this
network? How many components does it have? Is the network directed or
undirected? Do the nodes have attributes? Do the links have attributes?
Include your answers to these questions in your written report, but not
your code or its output.**

-   In R, the method to find the order of a network is to use the
    `vcount(g)` function, which returns the number of nodes in the
    network. When run on Star Wars Character Interaction Network, the
    function returns the value 111, which corresponds to the order of
    the network

    ```{r echo = FALSE, warning=FALSE, results='hide'}

    nO <- vcount(g) # Calculates the order of the network
    nO # variable storing the order of the network
    ```

**2. Using `Tree`, `Multidimensional Scaling`, and `Kamada-Kawai`
layouts separately, what do you observe when you visualize your network
? Describe your observations for each of these layouts. Use at most 25
words for each layout. Include your observations in your written report,
but not your code or its output.**

**- Plot using** `tree` **layout**

-   The `Tree` shows an interesting character hierarchy, which makes
    sense, as *Star Wars* follows the Skywalker family through
    generations. The network combined the interactions of Anakin
    Skywalker and Darth Vader into one `Darth Vader` node, making it's
    outdegree significantly higher than nodes.

    ```{r echo = FALSE, warning=FALSE, results='hide'}
    set.seed(THX)

    ggraph(g,layout = 'tree') +
      geom_edge_fan(edge_linetype = 1, color = 'blue', alpha = 0.25) +
      geom_node_point(color = node_colors, size = 1, alpha = 0.75) + 
      theme_graph(base_family = 'Helvetica') +
      labs(title = 'Star Wars Character Interactions',
           subtitle = 'Displayed using a tree Layout ')
    ```

**- Plot using** `Multidimensional Scaling` **layout**

-   The `mds`layout produced a visually appealing graph, but there is a
    large cluster of nodes in one region, making it difficult to extract
    information from the visualization.

    ```{r echo = FALSE, warning=FALSE, results='hide'}
    set.seed(THX)

    ggraph(g,layout = 'mds') +
      geom_edge_fan(edge_linetype = 1, color = 'blue', alpha = 0.25) +
      geom_node_point(color = node_colors, size = 1, alpha = 0.75) +
      theme_graph(base_family = 'Helvetica') + 
      labs(title = 'Star Wars Character Interactions',
           subtitle = 'Displayed using multidimensional scaling layout')
    ```

**- Plot using** `Kamada-Kawai` **layout**

-   The `kk` layout created a plot that was visually appealing, and
    highlighted potential hubs, as well as nodes on the periphery.

    ```{r echo = FALSE, warning=FALSE, results='hide'}
    set.seed(THX)

    ggraph(g, layout = 'kk') +
      geom_edge_fan(edge_linetype = 1, color = 'blue', alpha = 0.25) + 
      geom_node_point(color = node_colors, size = 1, alpha = 0.75) +
      theme_graph(base_family = 'Helvetica') +
      labs(title = 'Star Wars Character Interactions',
           subtitle = 'Displayed using Kamada-Kawai layout')
    ```

**3. Create a data.frame that contains eccentricity, betweenness,
degree, transitivity, eigen centrality, and hub score for your network.
You need to look up the documentation for the last two statistics in
Help. Note that both eigen centrality and hub score return three lists
and you need to access one of them. Make sure that you make the
appropriate changes to the code. Do not include your code or its output
in your report.**

-   **Statistics for the network, including** `eccentricity`,
    `betweenness`, `degree`, `Transitivity`, `Eigen Centrality`, and
    `Hub Score`.

    ```{r echo = FALSE, warning=FALSE, results='hide'}
    # [TODO]: Figure out the Eigen Centrality stuffs ####
    #Creates the dataframe df by binding different vectors created by calling each function 
    suppressMessages(df <- bind_cols(enframe(eccentricity(g, mode = 'out')),
                                            enframe(betweenness(g)),
                                            enframe(degree(g)),
                                            enframe(transitivity(g, type = c('local'))),
                                            enframe(eigen_centrality(g)$vector),
                                            enframe(hub_score(g)$vector)))
    # narrow the dataframe df to hold 5 values
    df <- df |> select(name...1, value...2, value...4, value...6, value...8, value...10, value...12) 

    # names the columns in the dataframe
    names(df) <- c('Names','eccentricity','betweenness','degree','clustering', 'Eigen Centrality','hub') 

    # prints first few rows of the dataframe
    head(df) 

    # prints the last few rows of the dataframe
    tail(df)

    # prints a 'glimpse' or summary of the dataframe
    glimpse(df) 
    ```

**4. Create a data.frame that includes the average degree, maximum
degree, diameter, average clustering coefficient, average betweenness,
average eigen centrality, and average hub score for your network.
Include the output in your report but not your code. Do you notice
anything interesting in this data.frame? Is this observation you are
making also true when you look at the data.frame in Question 3?**

-   **Metrics for the network, including** Avgerage Degree `[ADEG]`,
    `Max Degree`, `Proportion`, `Diameter`, `Radius`, Average clustering
    Coefficient `[ACC]`, Average betweenness `[ABT]`, Average Eigen
    Centrality `[AEC]`, and Average Hub Score `[AHS]`

    -   The Average Degree of the graph was 8, which was suprising, as
        the values in the `head()` and `tail()` of the dataframe
        fluctuated between 4-6.

    ```{r echo = FALSE, warning=FALSE, message=FALSE}
    sum_frame <- df |> 
      summarize("ADEG" = mean(degree),
                    # Maximum Outdegree, Largest Hub?
                    delta = max(degree), 
                    # Largest eccentricity? or amount of nodes in diameter
                    diam = max(eccentricity),
                    # Smallest eccentricity? or amount of nodes in radius?
                    radius = min(eccentricity),
                    # Average clustering Coefficient
                    'ACC' = mean(transitivity(g)),
                    # Average betweenness
                    'ABT' = mean(betweenness),
                    # Average Eigen Centrality
                    'AEC' = mean(`Eigen Centrality`),
                    # Average Hub Score
                    'AHS' = mean(`hub`))
    sum_frame
    ```

**5. Find the average distance in your network and create the plot for
the distribution of distance in this network. What do you notice about
this plot? Describe your observations in at most 50 words. Include your
answer and the plot in your report, but not your code or other
outputs.**

-   **Average Distance in the network**

    -   The average distance in the network is `2.610008`. This along
        with the plot reveals that the network is relatively dense,
        needing only 2.61 edges on average to connect two characters.

    ```{r echo = FALSE, warning=FALSE, results='hide'}
    avg <- mean_distance(g)  
    avg
    ```

    ```{r echo = FALSE, warning=FALSE, results='hide'}
    distance_table(g)

    D <- data.frame(1:length(distance_table(g)$res), 
                    distance_table(g)$res / sum(distance_table(g)$res))

    names(D) <- c('x', 'y')

    D |> 
      ggplot(aes(x = x, y = y)) + 
      geom_point() + 
      geom_line(aes(x = avg), color = 'blue') +
      labs(title = 'Distribution of Distance in the Star Wars Character Interaction Network') +
      labs(x = 'Distance', y = 'Density')
    ```

**6. Create the plot for the degree distribution in your network. What
do you notice about this plot? Describe your observations in at most 25
words. Include your answer and one plot in your report, but not your
code or other outputs.**

```{r echo = FALSE, warning=FALSE, results='hide'}
degrees <- degree(g)
# Create a data frame from the degrees
degree_data <- data.frame(degree = degrees)

# Create the histogram
ggplot(degree_data, aes(x = degree)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(title = "Degree Distribution of Character Interactions in Star Wars", x = "Degree", y = "Frequency")
```

**7. Based on Question 6, do you think that the degree distribution
follows a power law in your network? In other words, is this is a
scale-free network? If so, find its degree exponent.**

-   The previous plot of the degree distribution clearly shows that this
    network follows a power law distribution, as the plot is heavily
    right skewed.

**8. Create a plot (using dots rather histograms) containing both the
degree distribution for your network (use blue) and the Poisson
distribution with the correct parameter (use red). Include the plot in
your report, but not your code.**

-   The plot containing both the degree distribution for my network and
    the poisson distribution revealed that the degree distribution of
    the network follows a poisson-like pattern.

```{r echo = FALSE, warning=FALSE, results='hide'}
# Calculate mean degree for Poisson distribution
mean_degree <- mean(degrees)

# Create data frame for Poisson distribution
x <- c(0:max(degrees))

y <- dpois(x, mean(degrees))


df2 <- data.frame(x, y)

names(df2) <- c('x', 'y')

total <- sum(degrees)

df |> 
  group_by(degree) |> 
  mutate(x = degree,  
         y = n()/gorder(g)) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point(color = 'blue', size = .5) + 
  geom_point(data = df2, 
             aes(x = x, y = y), 
             color = 'red', 
             size = .5) + 
  labs(title = 'Dotplot of the Degree Distribution in the Star Wars Character Network (Blue)\nVersus that of a Poisson Distribution (Red)', x = 'Degree', y = TeX('$\\lambda'))
```

**9. Create a plot that visualizes the average of clustering
coefficients based on the degree in your network. Is the average
clustering coefficient matches what the Erdos-R ́enyi model predicts?
Include your answer and one plot in your report, but not your code or
other outputs.**

-   The plot that visualizes the average clustering coefficients based
    on the degree in my network shows that as the degree of a node
    increases, its clustering coefficient decreases. This is consistent
    with what the Erdos-Renyi model predicts. When compared against the
    random model, the graph has a higher average clustering coefficient,
    than what would be expected for a random graph.

```{r echo = FALSE, warning=FALSE, results='hide'}
C <- mean(df$clustering, na.rm = TRUE)

M <- mean(df$degree)

df |> 
  group_by(degree) |> 
  summarise(cc_deg = mean(clustering))

df |> 
  group_by(degree) |> 
  summarise(cc_deg = mean(clustering)) |> 
  ggplot(aes(x = degree, y = cc_deg)) + 
  geom_line(aes(y = C), linetype = "solid", color = "lightblue", size = 2, alpha =.75, lineend = "round") + 
  geom_line(aes(y = C), linetype = "solid", color = "#72bcd4", size = 1, alpha =1, lineend = "round") + 
  geom_line(aes(y = C), linetype = "solid", color = "white", size = .8, alpha =1, lineend = "round") + 
  geom_line(aes(y = M/gorder(g)), linetype = "solid", color = "#ff8080", size = 2, alpha =.75, lineend = "round") + 
  geom_line(aes(y = M/gorder(g)), linetype = "solid", color = "#ff4d4d", size = 1, alpha =1, lineend = "round") + 
  geom_line(aes(y = M/gorder(g)), linetype = "solid", color = "white", size = .8, alpha =1, lineend = "round") + 
  geom_point(na.rm = TRUE) + 
  scale_x_log10() + 
  scale_y_log10() + 
  labs(title =  'Relation Between Local Clustering Coefficient and Outdegree',
       subtitle = 'The blue line is the average local clustering coefficient; \nthe red line is the clustering coefficient predicted by the random model.') +
  labs(x = 'k', y = TeX('$C(k)$'))
```

**10. Create a plot that visualizes the relationship between betweenness
and degree in your network. What do you notice about this plot? Describe
your observations in at most 50 words. Include your answer and the plot
in your report, but not your code or other outputs.**

-   The plot of the relationship between betweenness and degree in my
    network revealed that as the degree of a node increases, so does
    it's betweenness centrality. This is expected as nodes with high
    betweenness, are often hubs, and hubs often have higher degrees.

```{r echo = FALSE, warning=FALSE, results='hide'}
df |> 
  filter(betweenness > 0) |> 
  ggplot(aes(x = degree, y = betweenness)) + 
  geom_point(na.rm = TRUE, size = 0.5, color = 'red') + 
  scale_y_log10() +
  labs(title = TeX('Relationship Between $\\log_{10}$ of Betweenness Centrality and Degree')) +
  labs(y = TeX('$\\log_{10}$(betweenness)'), x = 'Degree')
```

**11. Create a plot that visualizes the relationship between
eccentricity and degree in your network. What do you notice about this
plot? Describe your observations in at most 50 words. Include your
answer and the plot in your report, but not your code or other
outputs.**

-   The plot of eccentricity and degree reinforce that the network's
    degree distribution is close to the mean degree.

```{r echo = FALSE, warning=FALSE, results='hide'}
average_eccentricity <- mean(df$eccentricity, na.rm = TRUE)

df |> 
  ggplot(aes(x = degree, y = eccentricity)) + 
  geom_hline(yintercept = average_eccentricity, linetype = "solid", color = "#37e300", size = 2, alpha =.75, lineend = "round") +
  geom_hline(yintercept = average_eccentricity, linetype = "solid", color = "#00b30A", size = 1, lineend = "round") +
  geom_hline(yintercept = average_eccentricity, linetype = "solid", color = "white", size = .8, alpha = 1, lineend = "round") +
  geom_point(na.rm = TRUE, 
             size = .5, 
             color = 'red') +
  labs(title = 'Relationship Between Eccentricity and Degree',subtitle = 'Green represents the Average Eccentricity', x = 'Degree', y = 'Eccentricity')

```

**12. Create a plot that visualizes the relationship between clustering
and degree in your network. What do you notice about this plot? Describe
your observations in at most 50 words. Include your answer and the plot
in your report, but not your code or other outputs.**

-   The plot that visualizes the relationship between clustering and
    degree in my network reveals that it is not similar to the Random
    Model.

```{r echo = FALSE, warning=FALSE, results='hide'}
# Calculate the clustering coefficient for each node
node_cc <- transitivity(g, type = "local")

# Create a data frame to store the node IDs, degrees, and clustering coefficients
node_data <- data.frame(
  node = V(g)$name,
  degree = degree(g),
  clustering_coefficient = node_cc
)

# Calculate the average clustering coefficient for each degree
degree_avg_cc <- aggregate(clustering_coefficient ~ degree, data = node_data, FUN = mean)

# Plot the average clustering coefficient for each degree
ggplot()+
  geom_point(data = degree_avg_cc, aes(x = degree, y = clustering_coefficient),color='blue') + 
    labs(title = "Distribution of Clustering Coefficients",
         subtitle = "In the Star Wars Character Interaction Network",
         x = 'Degree',
         y = 'Clustering Coefficient')
```

**13. Create a plot that visualizes the relationship between hub score
and degree in your network. What do you notice about this plot? Describe
your observations in at most 50 words. Include your answer and the plot
in your report, but not your code or other outputs.**

-   The plot that visualizes the relationship between hub score and
    degree in my network revealed that as a node's degree increases, its
    hub score also increases, which is expected.

```{r echo = FALSE, warning=FALSE, results='hide'}
df |> 
  ggplot(aes(x = degree, y = hub)) + 
  geom_point(na.rm = TRUE, 
             size = 0.5, 
             color = 'red') +
  labs(title = 'Relationship Between Hub Score and Degree', x = 'Degree', y = 'Hub')
```
